Thesis Title

### 2022-03-25 chapters/introduction.tex
update

\section{The global impact of microbial pathogens} \label{sec:global_impact}

The Global Burden of Disease (GBD) 2019 study reported that microbial pathogens are responsible for more than 400 million years of life lost annually across the globe, a higher burden than either cancer or cardiovascular disease \citep{vos_global_2020}. 
In particular, lower respiratory infections, diarrhoeal diseases, HIV/AIDS and tuberculosis were amongst the five leading causes of global total years of life lost. 
More recently, the COVID-19 pandemic, declared as such by the World Health Organization (WHO) on 11 March 2020 after the emergence and global spread of the severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2), and as of January 2022, has caused more than 5.63 million deaths worldwide \citep{ritchie_coronavirus_2020}, making it one of the deadliest pandemics in history.
Coronavirus has been responsible for three of the eighteen major pandemics registered throughout modern history \citep{piret_pandemics_2021}, all occurring after the year 2000.
\textit{Yersinia pestis}, responsible for three pandemics of plague, \textit{Vibrio cholerae}, with seven cholera pandemics, and Influenza A virus, the causative agent of five flu pandemics, are responsible for the remaining, with Influenza being the only other pathogen with a pandemic registered after 2000. 
Recent decades have also witnessed the emergence of additional virulent pathogens, including the Ebola virus, West Nile virus, Dengue virus and Zika virus, particularly in lower-income countries.

@@

\subsection{Current standards for diagnostic in clinical microbiology} \label{ssec:current_standards}

The past few decades have seen a major revolution in the operation of microbial laboratories, driven by the development of molecular technologies and ways to make these accessible, namely amplification-based polymerase chain reaction (PCR), matrix-assisted laser desorption/ionisation - time of flight (MALDI-TOF) and DNA-microarray-based hybridisation technology. 
These are used in conjunction with traditional techniques such as microscopy, culture and serology, not fully replacing them.
Application of these methods differs by suspected infection type: bacterial, viral, fungal or parasitic. 
For the purpose of this dissertation work, we will be focusing on bacterial and viral infections.

\subsubsection{Bacterial infections} \label{sssec:bacterial}

@@

Choosing an appropriate bacterial typing technique for epidemiological studies depends on the resources available and the minimum intended resolution, ranging from DNA fingerprinting to multilocus sequence typing, Pulsed-field gel electrophoresis (PFGE) and sequence-based typing (see \secref{sec:genomics_approach}) \citep{allerberger_molecular_2012,foxman_choosing_2005}. 
DNA macrorestriction analysis by PFGE, which revolutionised precise separation of DNA fragments, became the most widely implemented DNA fingerprinting technique \citep{allerberger_molecular_2012}, becoming the golden standard for bacterial typing \citep{neoh_pulsed-field_2019}.

In the early 2000s, Multilocus sequence typing (MLST) was proposed as a portable, universal, and definitive method for characterising bacteria \citep{maiden_multilocus_2006}. 
Instead of enzyme restriction of bacteria DNA, separation of the restricted DNA bands using a PFGE chamber, followed by clonal assignment of bacteria based on banding patterns, MLST relies on the amplification through PCR sequences of internal fragments of housekeeping genes (usually 5 to 7), approximately 450-500 basepairs (bp) in size, followed by its the sequence, usually my Sanger methods (see \secref{ssec:1st_gen_seq}). 
For each house-keeping gene, the different sequences present within a bacterial species are assigned as distinct alleles and, for each isolate, the alleles at each of the (usually) seven loci define the allelic profile or sequence type \citep{larsen_multilocus_2012}. 
As with PFGE, different schemes, defining what house-keeping gene fragments are used, are available depending on the species. 
Unlike PFGE, the provision of freely accessible, curated databases of MLST nucleotide sequence data enables the direct comparison of bacterial isolates, providing the basis of a common language for bacterial typing \citep{maiden_multilocus_2006}. 
So far, MLST schemes for 115 bacterial organisms have been published and made freely available\footnote{\url{https://pubmlst.org/organisms}}, \cite{jolley_open-access_2018})

@@

\subsubsection{Viral infections} \label{sssec:viral}

@@

Direct detection methods negate the need for virus propagation, detecting the virus directly from the suspect source through nucleic acid and immunological methods. 
PCR and reverse transcription-PCR (RT-PCR) are widely applied methods for the detection of both DNA and RNA viruses, respectively, driven by increased awareness of the clinical value of, and demand for, prompt information about viral loads, viral sequence data, and potential antiviral resistance information \citep{cassedy_virus_2021}. 
Syndromic testing (see \secref{sssec:bacterial}) is now fully integrated into the standard testing practices of many clinical laboratories \citep{dien_bard_panels_2020}. 
Limitations of these assays include no detection of off-target pathogens, a lack of full susceptibility information, cost, and false-positive results. 
Real-time quantitative PCR (qPCR) remains the front line tool in aetiological diagnosis, measuring the production of the target amplicon throughout the reaction and providing quantitative results with high specificity and sensibility, albeit with a significant cost due to sophisticated apparatus despite high-throughput systems being widely established \citep{cassedy_virus_2021}.

@@

\subsection{Surveillance and infection prevention in public health} \label{ssec:survaillance}

@@

\section{A genomic approach to clinical microbiology} \label{sec:genomics_approach}

@@

\subsection{Twenty five years of microbial genome sequencing} \label{ssec:sequencing}

@@

\subsubsection{The first-generation of DNA sequencing} \label{ssec:1st_gen_seq}

@@

\subsubsection{The second-generation of DNA sequencing} \label{ssec:2nd_gen_seq}

@@

\paragraph{Sequencing by hybridisation} \label{sssec:2nd_gen_seq_hybrid} \mbox{}\\

@@

\paragraph{Sequencing by synthesis} \label{sssec:2nd_gen_seq_synth} \mbox{}\\

@@

\subsubsection{The third-generation of DNA sequencing} \label{ssec:3rd_gen_seq}

@@

\subsection{DNA sequencing in clinical diagnosis and surveillance} \label{ssec:sequencing_diagnosis}

Whole-genome sequencing (WGS) is becoming one of the most widely used applications of microbial genome sequencing. 
The major advantage of WGS is to yield all the available DNA information content on isolates in a single rapid step following culture (sequencing without culture will be discussed in the \secref{ssec:metagenomics}). 
In principle, after obtaining a pure culture, either bacterial (see \secref{sssec:bacterial}) or viral (see \secref{sssec:viral}), the data from sequencing contain all the information currently used for diagnostic and typing needs, and much more, thus opening the prospect for large-scale research into pathogen genotype-phenotype associations from routinely collected data \citep{didelot_transforming_2012}.
The cost of producing massive amounts of information requires a new framework with expert handling and processing of computer-driven genomic information, as well as capable computational infrastructures, but through this technology, researchers and clinicians can obtain the most comprehensive view of genomic information and associated biological implications, transforming clinical diagnostics and the detection and surveillance of outbreaks. \citep{cirulli_uncovering_2010, nature_reviews_genetics_genomic_2019, goodwin_coming_2016}.

@@

\subsubsection{Sequencing in the routine laboratory workflow} \label{sssec:sequencing_routine_lab}

@@

Although the second-generation DNA sequencing methods have shed light on fundamental aspects of microbial ecology and function, they suffer from issues associated with short read length (see \ref{ssec:2nd_gen_seq}) and cannot reliably reconstruct long repeats because of uncertainties in mapping read, even when paired-end sequencing is used. 
Third-generation sequencing methods (see \ref{ssec:3rd_gen_seq}) have become increasingly used in microbiology, although their accuracy and low throughput make it challenging to implement in a clinical diagnostic setting.

\subsubsection{Sequencing and genomic surveillance} \label{sssec:sequencing_genomic_survaillance}

@@

\subsection{From genomics to metagenomics} \label{ssec:metagenomics}

Despite the increasing adoption of DNA sequencing methods in clinical microbiology, the sequencing of genetic material from a pure culture requires \textit{a priori} knowledge of what to expect from a particular clinical sample or patient \citep{schuele_future_2021}. 
In most cases, this knowledge is enough to request the most appropriate test, such as multiplexed panels or specific culture media, but this is not always the case. 
In recent years, there has been a growing interest in using metagenomics to deliver culture-independent approaches to microbial ecology, surveillance and diagnosis (see Figure \ref{fig:figure4})\citep{loman_twenty_2015, loman_high-throughput_2012}.
Metagenomic DNA sequence allows detailed characterisation of pathogens in all kinds of samples originating from humans, animals, food and the environment, ligating the diagnostics to surveillance in a true "one health" fashion \citep{rossen__2018}. 
Unlike PCR or microarrays, it usually does not require primer or probe design, it can be easily multiplexed, and the specificity and selectivity of the sequencing can be adjusted computationally after acquiring the data \citep{dunne_next-generation_2012}.  
While most molecular assays target only a limited number of pathogens, metagenomic approaches characterise all DNA or RNA present in a sample, enabling analysis of the entire microbiome as well as the human host genome or transcriptome in patient samples \citep{chiu_clinical_2019}. 
Whether or not it can entirely replace routine microbiology depends on several conditions and future developments, both technological and computational (see \secref{sec:bioinformatics}).

@@

\subsubsection{Metataxonomics and Targeted Metagenomics} \label{sssec:metataxonomics}

@@

Because this approach is PCR-based, it suffers from the same issues described previously for conventional PCR, requiring primer design. 
Additionally, it must necessarily account for intragenomic variation between 16S gene copies. 
Microbial profiles generated using different primer pairs need independent validation of performance, and the comparison of data sets across V-regions using different databases might be misleading due to differences in nomenclature and varying precisions in classification, and specific but important taxa are not picked up by certain primer pairs (e.g., \textit{Bacteroidetes} is missed using primers 515F-944R) or due to the database used \citep{abellan-schneyder_primer_2021}. 
Furthermore, targeting of 16S variable regions with short-read sequencing platforms cannot achieve the taxonomic resolution afforded by sequencing the entire (~1500 bp) gene \citep{johnson_evaluation_2019}. 
The emergence of third generating sequencing technologies (see \secref{ssec:3rd_gen_seq}) allows for this limitation to be overcome but currently, only a fraction of the databases includes complete 16S rRNA sequences.

@@

\subsubsection{Shotgun Metagenomics} \label{sssec:shotgun_metagenomics}

@@

Clinical applications of shotgun metagenomics derive its roots from the use of microarrays (see \secref{ssec:current_standards}), where it was successfully applied in in-depth microbiome analysis of different sites in the human body, it was the emergence of second-generation sequencing technology and its high throughput of genomic data at a competitive price that made the sequencing of all genomic content, DNA and/or RNA) if a clinical sample a viable possibility for diagnostics (see \secref{ssec:2nd_gen_seq}) \citep{miller_basic_2009, palmer_rapid_2006, chiu_clinical_2019}. The first reported case that demonstrated the utility of shotgun metagenomics was in 2014 with the clinical diagnosis of neuroleptospirosis in a 14-year-old immunodeficient and critically ill boy with meningoencephalitis by \cite{wilson_actionable_2014}, prompting appropriate targeted antibiotic treatment and eventual recovery of the patient. In this case, traditional methods, including an invasive brain biopsy, failed to provide answers, until the shotgun sequencing of cerebrospinal fluid identified 475 of 3,063,784 sequence reads (0.016\%) corresponding to leptospira, for which clinical assays were negative due to its very low abundance. Ever since many other reports of successful application of shotgun metagenomics in clinical metagenomics have been reported. but all in edge cases where traditional diagnostic methods have failed or as proof-of-concept \citep{couto_critical_2018, vijayvargiya_application_2019, sanabria_shotgun-metagenomics_2020, hirakata_application_2021}.

@@

In addition to negative controls, already a common practice in any sequencing assay and in particular in metataxonomics (see \secref{sssec:metataxonomics}), positive controls can be a way to circumvent the lack of golden standards, either through the spike of the samples with a known amount of a specific DNA/RNA or though the sequencing of samples with known composition and abundance. Well-characterised reference standards and controls are needed to ensure shotgun metagenomics assay quality and stability over time \citep{chiu_clinical_2019, mcintyre_comprehensive_2017}. Most available metagenomic reference materials are highly tailored to a specific application. For example, the ZymoBIOMICS Microbial Community Standard\footnote{\url{https://www.zymoresearch.com/collections/zymobiomics-microbial-community-standards}} is the first commercially available standard for microbiomics and metagenomics studies, providing mock a mock community with defined composition and abundance consisting of Gram-positive, Gram-negative and yeast. It is useful to determine the limit of detection of an assay, and the effectiveness and biases of a given protocol. Standards with a more limited spectrum of organisms are also available, such as the National Institute of Standards and Technology (NIST)\footnote{\url{https://www.nist.gov/}} reference materials for mixed microbial DNA detection, which contain only bacteria. Thus, these materials may not apply to untargeted shotgun metagenomics analyses.

\section{The role of bioinformatics} \label{sec:bioinformatics}

As stated previously (see \secref{sec:genomics_approach} and \secref{ssec:metagenomics}), one of the biggest challenges when dealing with genomic, and in particular metagenomic, data is the lack of golden standards. This is also applicable to the bioinformatic analysis, required due to the amount of data produced by genomic sequencing technologies. This is currently one of the bottlenecks in the deployment of sequencing technology in clinical microbiology as there's no standard in how to deal with the increasing amount of data produced in a fit-for-purpose manner \citep{carrico_primer_2018}.

@@

\subsection{The FASTQ file} \label{ssec:fastq}

In all sequencing technologies (see \secref{ssec:sequencing}), many copies of the source DNA are randomly fragmented and sequenced. To these sequences, we refer to as reads. In the case of second-generation sequencing (see \secref{ssec:2nd_gen_seq}), one or both ends of the fragment can be sequenced. If a fragment is sequenced from one end, we refer to it as single-end sequencing. If a fragment is sequenced on both ends, spanning the entire fragment, it is called paired-end sequencing.

@@

\subsubsection{FASTQ file simulation} \label{ssec:fastq_sim}

@@

\subsubsection{FASTQ quality assessment and quality control} \label{ssec:fastq_quality}

@@

\subsection{Direct taxonomic assignment and characterisation} \label{ssec:taxonomic_assignment}

@@

These tools are also extremely useful to assess the contents of a metagenomic sample. Alternatives such as Midas \citep{nayfach_integrated_2016}, Kaiju, \citep{menzel_fast_2016}, and MetaPhlAn2 \citep{truong_metaphlan2_2015} offer the same analysis as Kraken and Bracken using different algorithms, and with the disadvantage that they come prepackaged with their own databases, without the option to create a tailored database, limiting their applicability. Kaiju differs from the other tools by using a protein reference database, instead of nucleotide, but no pre-built version is available, requiring significant resources to build and index the database pre-use. The long-read data of third-generation sequencing technologies (see \secref{ssec:3rd_gen_seq}) can be treated as single-end reads, and all tools mentioned accommodate the classification of single-end files.

\subsection{From reads to genomes} \label{ssec:reads_2_genomes}

Due to the limitations of current sequencing technologies (see \secref{ssec:sequencing}), the order of the reads produced by these machines cannot be preserved. 
Therefore, to obtain the true original genomic sequence the process of "genome assembly" has to occur. 
The term "draft genome" is commonly used because these sequencing technologies do not generate a single closed genome, particularly short-read such as in second generation sequencing (see \secref{ssec:2nd_gen_seq}) which need to be assembled into usually a series of sequences (contigs) that may cover up to 95\% to 99\% of the strain genome \citep{carrico_primer_2018}. 
Long-read technologies (see \secref{ssec:3rd_gen_seq}) allow for this value to reach 100\%, effectively producing closed, complete genomes, notwithstanding that this value can sometimes overcome the 100\% due to overlap \citep{wick_benchmarking_2021}.

@@

\paragraph{Overlap, Layout and Consensus assembly} \label{sssec:OLC_assembly} \mbox\\

First generation sequencing technology (see \secref{ssec:1st_gen_seq}) produces far fewer reads than second generation sequencing technology (see \secref{ssec:2nd_gen_seq}, but individual reads are longer (500–1000 bp). 
Assembly of Sanger data usually uses overlap-layout consensus (OLC) approaches, in which:

@@

These type of assemblers were very popular in the early 2010s, with assemblers such as Celera\footnote{\url{https://www.cbcb.umd.edu/software/celera-assembler}},  Genovo\footnote{\url{https://cs.stanford.edu/genovo }}, xGenovo\footnote{\url{http://xgenovo.dna.bio.keio.ac.jp/}} and BBAP\footnote{\url{http://homepage.ntu.edu.tw/~youylin/BBAP.html}} having been widely used \citep{myers_whole-genome_2000, hutchison_genovo_2010, afiahayati_extended_2013, lin_novo_2017}.
With the emergence of third-generation sequencing (see \secref{ssec:3rd_gen_seq}), OLC assemblers have been increasingly developed and adopted by the community to assembly long-read data. 
In the latest years, ra\footnote{\url{https://github.com/lbcb-sci/ra}}, raven\footnote{\url{https://github.com/lbcb-sci/raven}} and canu\footnote{\url{https://github.com/marbl/canu}}, the latter being a a fork of the Celera Assembler, have become staples in the community, showing good reliability and amassing over 3000 citations \citep{vaser_yet_2019, koren_canu_2017, wick_benchmarking_2021}.  %missing raven

\paragraph{De Bruijn graph assembly} \label{sssec:dbg_assembly} \mbox{}\\

@@

\subsection{Reproducibility and transparency} \label{ssec:reproducibility}

@@

\section{Bioinformatic Analysis for Metagenomics} \label{sec:metagenomics_bioinfo}

As mentioned previously (see \secref{ssec:metagenomics}, Metagenomic shotgun sequencing circumvents the need for cultivation and, compared with metataxonomics, avoids biases from primer choice, enables the detection of organisms across all domains of life and \textit{de novo} assembly of genomes and functional genome analyses. However, highly uneven sequencing depth of different organisms and low depth of coverage per species are drawbacks that limit taxa

\subsubsection{Metataxonomics} \label{ssec:metataxonomics_bioinfo}

>>>>

\section{The global impact of microbial pathogens} \label{sec:_intro_global_impact}

The Global Burden of Disease (GBD) 2019 study reported that microbial pathogens are responsible for more than 400 million years of life lost annually across the globe, a higher burden than either cancer or cardiovascular disease \citep{vos_global_2020}. 
In particular, lower respiratory infections, diarrhoeal diseases, HIV/AIDS and tuberculosis were amongst the five leading causes of global total years of life lost. 
More recently, the COVID-19 pandemic, declared as such by the World Health Organization (WHO) on 11 March 2020 after the emergence and global spread of the severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2), has caused more than 5 million deaths worldwide \citep{ritchie_coronavirus_2020}, making it one of the deadliest pandemics in history.
Coronavirus has been responsible for three of the eighteen major pandemics registered throughout modern history \citep{piret_pandemics_2021}, all occurring after the year 2000.
\textit{Yersinia pestis}, responsible for three pandemics of plague, \textit{Vibrio cholerae}, with seven cholera pandemics, and Influenza A virus, the causative agent of five flu pandemics, are responsible for the remaining, with Influenza being the only other pathogen with a pandemic registered after 2000. 
Recent decades have also witnessed the emergence of additional virulent pathogens, including the Ebola virus, West Nile virus, Dengue virus and Zika virus, particularly in lower-income countries.

@@

\subsection{Current standards for diagnostic in clinical microbiology} \label{ssec:_intro_current_standards}

The past few decades have seen a major revolution in the operation of microbial laboratories, driven by the development of molecular technologies and ways to make these accessible, namely amplification-based polymerase chain reaction (PCR), matrix-assisted laser desorption/ionisation - time of flight (MALDI-TOF) and DNA-microarray-based hybridisation technology. 
These are used in conjunction with traditional techniques such as microscopy, culture and serology.
Application of these methods differs by suspected infection type: bacterial, viral, fungal or parasitic. 
For the purpose of this dissertation work, we will be focusing on bacterial (see Section \ref{sssec:_intro_bacterial} and viral infections (see Section \ref{sssec:_intro_viral}.

\subsubsection{Bacterial infections} \label{sssec:_intro_bacterial}

@@

Choosing an appropriate bacterial typing technique for epidemiological studies depends on the resources available and the minimum intended resolution, ranging from DNA fingerprinting to multilocus sequence typing, Pulsed-field gel electrophoresis (PFGE) and sequence-based typing (see \secref{sec:_intro_genomics_approach}) \citep{allerberger_molecular_2012,foxman_choosing_2005}. 
DNA macrorestriction analysis by PFGE, which revolutionised precise separation of DNA fragments, became the most widely implemented DNA fingerprinting technique \citep{allerberger_molecular_2012}, becoming the golden standard for bacterial typing \citep{neoh_pulsed-field_2019}.

In the early 2000s, Multilocus sequence typing (MLST) was proposed as a portable, universal, and definitive method for characterising bacteria \citep{maiden_multilocus_2006}. 
Instead of enzyme restriction of bacteria DNA, separation of the restricted DNA bands using a PFGE chamber, followed by clonal assignment of bacteria based on banding patterns, MLST relies on the amplification through PCR sequences of internal fragments of housekeeping genes (usually 5 to 7), approximately 450-500 basepairs (bp) in size, followed by its the sequence, usually my Sanger methods (see \secref{ssec:_intro_1st_gen_seq}). 
For each house-keeping gene, the different sequences present within a bacterial species are assigned as distinct alleles and, for each isolate, the alleles at each of the (usually) seven loci define the allelic profile or sequence type \citep{larsen_multilocus_2012}. 
As with PFGE, different schemes, defining what house-keeping gene fragments are used, are available depending on the species. 
Unlike PFGE, the provision of freely accessible, curated databases of MLST nucleotide sequence data enables the direct comparison of bacterial isolates, providing the basis of a common language for bacterial typing \citep{maiden_multilocus_2006}. 
So far, MLST schemes for 115 bacterial organisms have been published and made freely available\footnote{\url{https://pubmlst.org/organisms}}, \cite{jolley_open-access_2018})

@@

\subsubsection{Viral infections} \label{sssec:_intro_viral}

@@

Direct detection methods negate the need for virus propagation, detecting the virus directly from the suspect source through nucleic acid and immunological methods. 
PCR and reverse transcription-PCR (RT-PCR) are widely applied methods for the detection of both DNA and RNA viruses, respectively, driven by increased awareness of the clinical value of, and demand for, prompt information about viral loads, viral sequence data, and potential antiviral resistance information \citep{cassedy_virus_2021}. 
Syndromic testing (see \secref{sssec:_intro_bacterial}) is now fully integrated into the standard testing practices of many clinical laboratories \citep{dien_bard_panels_2020}. 
Limitations of these assays include no detection of off-target pathogens, a lack of full susceptibility information, cost, and false-positive results. 
Real-time quantitative PCR (qPCR) remains the front line tool in aetiological diagnosis, measuring the production of the target amplicon throughout the reaction and providing quantitative results with high specificity and sensibility, albeit with a significant cost due to sophisticated apparatus despite high-throughput systems being widely established \citep{cassedy_virus_2021}.

@@

\subsection{Surveillance and infection prevention in public health} \label{ssec:_intro_survaillance}

@@

\section{A genomic approach to clinical microbiology} \label{sec:_intro_genomics_approach}

@@

\subsection{Twenty five years of microbial genome sequencing} \label{ssec:_intro_sequencing}

@@

\subsubsection{The first-generation of DNA sequencing} \label{ssec:_intro_1st_gen_seq}

@@

\subsubsection{The second-generation of DNA sequencing} \label{ssec:_intro_2nd_gen_seq}

@@

\paragraph{Sequencing by hybridisation} \label{sssec:_intro_2nd_gen_seq_hybrid} \mbox{}\\

@@

\paragraph{Sequencing by synthesis} \label{sssec:_intro_2nd_gen_seq_synth} \mbox{}\\

@@

\subsubsection{The third-generation of DNA sequencing} \label{ssec:_intro_3rd_gen_seq}

@@

\subsection{DNA sequencing in clinical diagnosis and surveillance} \label{ssec:_intro_sequencing_diagnosis}

Whole-genome sequencing (WGS) is becoming one of the most widely used applications of microbial genome sequencing. 
The major advantage of WGS is to yield all the available DNA information content on isolates in a single rapid step following culture (sequencing without culture will be discussed in the \secref{ssec:_intro_metagenomics}). 
In principle, after obtaining a pure culture, either bacterial (see \secref{sssec:_intro_bacterial}) or viral (see \secref{sssec:_intro_viral}), the data from sequencing contain all the information currently used for diagnostic and typing needs, and much more, thus opening the prospect for large-scale research into pathogen genotype-phenotype associations from routinely collected data \citep{didelot_transforming_2012}.
The cost of producing massive amounts of information requires a new framework with expert handling and processing of computer-driven genomic information, as well as capable computational infrastructures, but through this technology, researchers and clinicians can obtain the most comprehensive view of genomic information and associated biological implications, transforming clinical diagnostics and the detection and surveillance of outbreaks. \citep{cirulli_uncovering_2010, nature_reviews_genetics_genomic_2019, goodwin_coming_2016}.

@@

\subsubsection{Sequencing in the routine laboratory workflow} \label{sssec:_intro_sequencing_routine_lab}

@@

Although the second-generation DNA sequencing methods have shed light on fundamental aspects of microbial ecology and function, they suffer from issues associated with short read length (see \ref{ssec:_intro_2nd_gen_seq}) and cannot reliably reconstruct long repeats because of uncertainties in mapping read, even when paired-end sequencing is used. 
Third-generation sequencing methods (see \ref{ssec:_intro_3rd_gen_seq}) have become increasingly used in microbiology, although their accuracy and low throughput make it challenging to implement in a clinical diagnostic setting.

\subsubsection{Sequencing and genomic surveillance} \label{sssec:_intro_sequencing_genomic_survaillance}

@@

\subsection{From genomics to metagenomics} \label{ssec:_intro_metagenomics}

Despite the increasing adoption of DNA sequencing methods in clinical microbiology, the sequencing of genetic material from a pure culture requires \textit{a priori} knowledge of what to expect from a particular clinical sample or patient \citep{schuele_future_2021}. 
In most cases, this knowledge is enough to request the most appropriate test, such as multiplexed panels or specific culture media, but this is not always the case. 
In recent years, there has been a growing interest in using metagenomics to deliver culture-independent approaches to microbial ecology, surveillance and diagnosis (see Figure \ref{fig:figure4})\citep{loman_twenty_2015, loman_high-throughput_2012}.
Metagenomic DNA sequence allows detailed characterisation of pathogens in all kinds of samples originating from humans, animals, food and the environment, ligating the diagnostics to surveillance in a true "one health" fashion \citep{rossen__2018}. 
Unlike PCR or microarrays, it usually does not require primer or probe design, it can be easily multiplexed, and the specificity and selectivity of the sequencing can be adjusted computationally after acquiring the data \citep{dunne_next-generation_2012}.  
While most molecular assays target only a limited number of pathogens, metagenomic approaches characterise all DNA or RNA present in a sample, enabling analysis of the entire microbiome as well as the human host genome or transcriptome in patient samples \citep{chiu_clinical_2019}. 
Whether or not it can entirely replace routine microbiology depends on several conditions and future developments, both technological and computational (see \secref{sec:_intro_bioinformatics}).

@@

\subsubsection{Metataxonomics and Targeted Metagenomics} \label{sssec:_intro_metataxonomics}

@@

Because this approach is PCR-based, it suffers from the same issues described previously for conventional PCR, requiring primer design. 
Additionally, it must necessarily account for intragenomic variation between 16S gene copies. 
Microbial profiles generated using different primer pairs need independent validation of performance, and the comparison of data sets across V-regions using different databases might be misleading due to differences in nomenclature and varying precisions in classification, and specific but important taxa are not picked up by certain primer pairs (e.g., \textit{Bacteroidetes} is missed using primers 515F-944R) or due to the database used \citep{abellan-schneyder_primer_2021}. 
Furthermore, targeting of 16S variable regions with short-read sequencing platforms cannot achieve the taxonomic resolution afforded by sequencing the entire (~1500 bp) gene \citep{johnson_evaluation_2019}. 
The emergence of third generating sequencing technologies (see \secref{ssec:_intro_3rd_gen_seq}) allows for this limitation to be overcome but currently, only a fraction of the databases includes complete 16S rRNA sequences.

@@

\subsubsection{Shotgun Metagenomics} \label{sssec:_intro_shotgun_metagenomics}

@@

Clinical applications of shotgun metagenomics derive its roots from the use of microarrays (see \secref{ssec:_intro_current_standards}), where it was successfully applied in in-depth microbiome analysis of different sites in the human body, it was the emergence of second-generation sequencing technology and its high throughput of genomic data at a competitive price that made the sequencing of all genomic content, DNA and/or RNA) if a clinical sample a viable possibility for diagnostics (see \secref{ssec:_intro_2nd_gen_seq}) \citep{miller_basic_2009, palmer_rapid_2006, chiu_clinical_2019}. The first reported case that demonstrated the utility of shotgun metagenomics was in 2014 with the clinical diagnosis of neuroleptospirosis in a 14-year-old immunodeficient and critically ill boy with meningoencephalitis by \cite{wilson_actionable_2014}, prompting appropriate targeted antibiotic treatment and eventual recovery of the patient. In this case, traditional methods, including an invasive brain biopsy, failed to provide answers, until the shotgun sequencing of cerebrospinal fluid identified 475 of 3,063,784 sequence reads (0.016\%) corresponding to leptospira, for which clinical assays were negative due to its very low abundance. Ever since many other reports of successful application of shotgun metagenomics in clinical metagenomics have been reported. but all in edge cases where traditional diagnostic methods have failed or as proof-of-concept \citep{couto_critical_2018, vijayvargiya_application_2019, sanabria_shotgun-metagenomics_2020, hirakata_application_2021}.

@@

In addition to negative controls, already a common practice in any sequencing assay and in particular in metataxonomics (see \secref{sssec:_intro_metataxonomics}), positive controls can be a way to circumvent the lack of golden standards, either through the spike of the samples with a known amount of a specific DNA/RNA or though the sequencing of samples with known composition and abundance. Well-characterised reference standards and controls are needed to ensure shotgun metagenomics assay quality and stability over time \citep{chiu_clinical_2019, mcintyre_comprehensive_2017}. Most available metagenomic reference materials are highly tailored to a specific application. For example, the ZymoBIOMICS Microbial Community Standard\footnote{\url{https://www.zymoresearch.com/collections/zymobiomics-microbial-community-standards}} is the first commercially available standard for microbiomics and metagenomics studies, providing mock a mock community with defined composition and abundance consisting of Gram-positive, Gram-negative and yeast. It is useful to determine the limit of detection of an assay, and the effectiveness and biases of a given protocol. Standards with a more limited spectrum of organisms are also available, such as the National Institute of Standards and Technology (NIST)\footnote{\url{https://www.nist.gov/}} reference materials for mixed microbial DNA detection, which contain only bacteria. Thus, these materials may not apply to untargeted shotgun metagenomics analyses.

\section{The role of bioinformatics} \label{sec:_intro_bioinformatics}

As stated previously (see \secref{sec:_intro_genomics_approach} and \secref{ssec:_intro_metagenomics}), one of the biggest challenges when dealing with genomic, and in particular metagenomic, data is the lack of golden standards. This is also applicable to the bioinformatic analysis, required due to the amount of data produced by genomic sequencing technologies. This is currently one of the bottlenecks in the deployment of sequencing technology in clinical microbiology as there's no standard in how to deal with the increasing amount of data produced in a fit-for-purpose manner \citep{carrico_primer_2018}.

@@

\subsection{The FASTQ file} \label{ssec:_intro_fastq}

In all sequencing technologies (see \secref{ssec:_intro_sequencing}), many copies of the source DNA are randomly fragmented and sequenced. To these sequences, we refer to as reads. In the case of second-generation sequencing (see \secref{ssec:_intro_2nd_gen_seq}), one or both ends of the fragment can be sequenced. If a fragment is sequenced from one end, we refer to it as single-end sequencing. If a fragment is sequenced on both ends, spanning the entire fragment, it is called paired-end sequencing.

@@

\subsubsection{FASTQ file simulation} \label{ssec:_intro_fastq_sim}

@@

\subsubsection{FASTQ quality assessment and quality control} \label{ssec:_intro_fastq_quality}

@@

\subsection{Direct taxonomic assignment and characterisation} \label{ssec:_intro_taxonomic_assignment}

@@

These tools are also extremely useful to assess the contents of a metagenomic sample. Alternatives such as Midas \citep{nayfach_integrated_2016}, Kaiju, \citep{menzel_fast_2016}, and MetaPhlAn2 \citep{truong_metaphlan2_2015} offer the same analysis as Kraken and Bracken using different algorithms, and with the disadvantage that they come prepackaged with their own databases, without the option to create a tailored database, limiting their applicability. Kaiju differs from the other tools by using a protein reference database, instead of nucleotide, but no pre-built version is available, requiring significant resources to build and index the database pre-use. The long-read data of third-generation sequencing technologies (see \secref{ssec:_intro_3rd_gen_seq}) can be treated as single-end reads, and all tools mentioned accommodate the classification of single-end files.

\subsection{From reads to genomes} \label{ssec:_intro_reads_2_genomes}

Due to the limitations of current sequencing technologies (see \secref{ssec:_intro_sequencing}), the order of the reads produced by these machines cannot be preserved. 
Therefore, to obtain the true original genomic sequence the process of "genome assembly" has to occur. 
The term "draft genome" is commonly used because these sequencing technologies do not generate a single closed genome, particularly short-read such as in second generation sequencing (see \secref{ssec:_intro_2nd_gen_seq}) which need to be assembled into usually a series of sequences (contigs) that may cover up to 95\% to 99\% of the strain genome \citep{carrico_primer_2018}. 
Long-read technologies (see \secref{ssec:_intro_3rd_gen_seq}) allow for this value to reach 100\%, effectively producing closed, complete genomes, notwithstanding that this value can sometimes overcome the 100\% due to overlap \citep{wick_benchmarking_2021}.

@@

\paragraph{Overlap, Layout and Consensus assembly} \label{sssec:_intro_OLC_assembly} \mbox\\

First generation sequencing technology (see \secref{ssec:_intro_1st_gen_seq}) produces far fewer reads than second generation sequencing technology (see \secref{ssec:_intro_2nd_gen_seq}, but individual reads are longer (500–1000 bp). 
Assembly of Sanger data usually uses overlap-layout consensus (OLC) approaches, in which:

@@

These type of assemblers were very popular in the early 2010s, with assemblers such as Celera\footnote{\url{https://www.cbcb.umd.edu/software/celera-assembler}},  Genovo\footnote{\url{https://cs.stanford.edu/genovo }}, xGenovo\footnote{\url{http://xgenovo.dna.bio.keio.ac.jp/}} and BBAP\footnote{\url{http://homepage.ntu.edu.tw/~youylin/BBAP.html}} having been widely used \citep{myers_whole-genome_2000, hutchison_genovo_2010, afiahayati_extended_2013, lin_novo_2017}.
With the emergence of third-generation sequencing (see \secref{ssec:_intro_3rd_gen_seq}), OLC assemblers have been increasingly developed and adopted by the community to assembly long-read data. 
In the latest years, ra\footnote{\url{https://github.com/lbcb-sci/ra}}, raven\footnote{\url{https://github.com/lbcb-sci/raven}} and canu\footnote{\url{https://github.com/marbl/canu}}, the latter being a a fork of the Celera Assembler, have become staples in the community, showing good reliability and amassing over 3000 citations \citep{vaser_yet_2019, koren_canu_2017, wick_benchmarking_2021}.  %missing raven

\paragraph{De Bruijn graph assembly} \label{sssec:_intro_dbg_assembly} \mbox{}\\

@@

\subsection{Reproducibility and transparency} \label{ssec:_intro_reproducibility}

@@

\section{Bioinformatic Analysis for Metagenomics} \label{sec:_intro_metagenomics_bioinfo}

As mentioned previously (see \secref{ssec:_intro_metagenomics}, Metagenomic shotgun sequencing circumvents the need for cultivation and, compared with metataxonomics, avoids biases from primer choice, enables the detection of organisms across all domains of life and \textit{de novo} assembly of genomes and functional genome analyses. However, highly uneven sequencing depth of different organisms and low depth of coverage per species are drawbacks that limit taxa

\subsubsection{Metataxonomics} \label{ssec:_intro_metataxonomics_bioinfo}
